---
title: "More visualizations with ggplot2"
author: "Bastola"
date: "1/12/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE,eval=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE, 
                      comment=NA, 
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align='center')


library(dplyr)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(dslabs)
library(datasauRus)
```

## Your Turn 1

The data frame in this exercise is called datasaurus_dozen and it's in the datasauRus package. This single data frame contains 13 datasets.

```{r, echo=FALSE}
# Use single dataset called 'dino'
dino_data <- datasaurus_dozen %>%
   filter(dataset == "dino")
```

a. Generate a scatter plot of the variables y and x in the dataset `dino_data` using `aes()` mapping function.

```{r}
ggplot(data = ____, mapping = aes(x = __, y = __)) +
  geom_point()
```

b. You _must_ remember to put the aesthetic mappings in the `aes()` function! What happens if you forget? 

```{r}
# Add a layer and see what happens
ggplot(data = ___ , x = ___ , y = __)
```

*Answer:*



c. The aesthetic mappings can be specified in the geom layer if you prefer, instead of the main `ggplot()` call. Give it a try:

```{r}
# Rebuild the scatterplot with your aesthetic mapping in the geom layer
ggplot(data = dino_data) +
   geom_point(aes(x = __, y = __)) 
```


\vspace*{1in}


# Your turn 2

Work with the same dataset `datasaurus_dozen` from the `datasauRus` package. This single data frame contains 13 datasets.

a. Let's `filter` 4 datasets called `dino`, `star`, `x_shape`, and `away`. Make a multipanel scatter plot usinf these 4 datasets.


```{r, echo=FALSE}
library(datasauRus)
dino_data_four <- datasaurus_dozen %>%
   filter(dataset == "dino" | dataset == "star" | dataset == "x_shape"| dataset == "away")
```


```{r}
ggplot(dino_data_four, aes(x = x, y = y, color = ______)) +
  geom_point() +
  facet_wrap(~ ______, ncol = 2) +
  theme(legend.position = "none")
```


b. Could you make a multi-panel plot including all 13 datasets. What is your favorite shape?

```{r}
ggplot(___________, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~ dataset, ncol = 4) +
  theme(legend.position = "none")

```

\vspace*{1in}


## Your Turn 3

- The `mpg` data set is loaded with the `tidyverse`. Run `?mpg` for more info. 

a. Create this bar chart of vehicle `class`.

```{r echo=FALSE, fig.height = 4, fig.wigth = 7, fig.align='center'}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = ______, fill = ______))
```


\vspace*{1in}

# Your Turn 4

The 2018 Gapminder data set contains information on income (GDP adjusted for difference in purchasing power), life expectancy (the average number of years a newborn is expected to live if mortality patterns remain constant), and population for 193 countries in 2018. 

```{r}
gap_dat <-read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/gapminder2018.csv")
gapminder <- gap_dat %>% filter(year == 2018)
glimpse(gapminder)
```

a. Build the histogram of `life expectancy` using `geom_histogram()`. Don't hesitate to look at the ggplot2 cheat sheet for help!

```{r}
# Fill in the blanks
ggplot(gapminder, mapping = aes(__________)) +
    geom_histogram()
```


b. What have you learned about the distribution of life expectancy in 2018?

*Answer:* 


c. To change the number of bins, to say 50, add the *argument* `bins = 50` to `geom_histogram()`. Note: this is **not** an aesthetic mapping.

```{r}
# Fill in the blanks
ggplot(gapminder, mapping = aes(life_expectancy)) +
    geom_histogram(_____)
```

d. Instead of a histogram, let's create a kernel density plot. To do this, substitute `geom_density()` into your code for question
4c.

```{r}
ggplot(gapminder, mapping = aes(life_expectancy)) +
  ____________
```


Next, let's make _boxplots_ of the population for each region (based on the `four_regions` column).

d. Use `geom_boxplot()` to create _box plots_ of the population for each region. What do you learn?

```{r}
# Fill in the blanks
ggplot(gapminder, mapping = aes(x = four_regions, y = population))
+ _________
```


*Answer:* 


e. The boxplots in question (d) are hard to read due to the _scale_ of the y-axis. Add the layer `scale_y_log10()` to apply a logarithmic transformation to the y-axis. What do you learn?

```{r}
ggplot(gapminder, mapping = aes(x = four_regions, y = population)) +
  geom_boxplot() +
  ____________
```

*Answer:*

f. A violin plot is a kernel density on its side, made symmetric. Change your code from question 4e to use `geom_violin()`. Which plot do you prefer, boxplots or violin plots? Why?

```{r}
ggplot(gapminder, mapping = aes(x = four_regions, y = population)) +
  _____________ +
  scale_y_log10()
```

How many countries are in each region? Let's use graphics to find out!


g. Make a bar chart of the number of countries in each continent using `geom_bar()`. 

```{r}
# Fill in the blanks
ggplot(gapminder, mapping = aes(x = four_regions)) +
  geom_bar()
```


*Answer:* 


h. Notice the y-axis. Oddly, `ggplot2` doesn't make it obvious how to change to proportion. Try adding a `y` aesthetic: `y = ..count../sum(..count..)`.

```{r}
ggplot(gapminder, mapping = aes(x = four_regions, y = ..count../sum(..count..))) +
  geom_bar()
```






